Percussive Sound Generation with Virtual Listeners and Modular Synthesizers

### 2020-08-16 templates_aimusic2020/paper.tex
Updates from Overleaf

\begin{figure}[h]

\centering
\includegraphics[width=1\linewidth,angle =-90 ]{images/periodic_function_decimation.png}
\caption{Reconstruction of a periodic, continuous wave at three different sampling levels. Higher sampling rates result in higher reconstruction accuracy. } % AH: this is linear interpolation many sound cards will convert the last signal back into the original wave just fine
\label{fig_sampling_rate}
\end{figure}

@@

% https://en.wikipedia.org/wiki/Frequency_domain#/media/File:Fourier_transform_time_and_frequency_domains_(small).gif

\section{Digital Audio Synthesis}
\label{sec_digital_synthesis}
The phenomena of sound at intensities we commonly encounter can be described as a product of a \textit{linear system} of functions~\cite{cook1999chap4}. A linear system is a system where the transformation of overlapping inputs is equal to the sum of the separately transformed inputs~\cite{lyons2004understandingChap1,cook1999chap4}. In a linear system $\mathcal{S}$ with valid inputs and outputs $x(i)$ and $y(i)$, if:

\begin{equation}
 \mathcal{S}(x(i_1)) \xrightarrow{generates} y(i_1)
\end{equation}
\begin{center}
    and
\end{center}
\begin{equation}
\mathcal{S}(x(i_2)) \xrightarrow{generates}y(i_2)
\end{equation}

The output of the system given both inputs is the sum of the individual outputs, or:
\begin{equation}
 \mathcal{S}(x(i_1)+x(i_2)) \xrightarrow{generates} y(i_1)+y(i_2) 
\end{equation}
\\
This concept has important implications digital audio creation and analysis. Simple tones can be combined to create complex sounds, and complex sounds can be broken down for easier analysis~\cite{lyons2004understandingChap1}. It also allows systems based on experiments with simple sine waves to remain relevant in complex sound domains~\cite{cook1999chap4}.  \\\\

[graph of addition of two sine waves making a complex sine wave ]
\\

Various sound synthesis techniques have been developed via treatment of sound as a sequence of values. Linear systems are commonly used in creation of musical tones, while non-linear systems are used for introduction of distortion and noise where needed. In their taxonomy of digital synthesis techniques, Smith defines four families of algorithms. Algorithms that process and modulate existing sounds (e.g granular synthesis, wavelets), Spectral models that aim to create a particular spectrum of sound (e.g additive, subtractive), physical models which emulate the physics of real instruments (discussed briefly in Section ~\ref{sec_tools_disposal}) and abstract models (e.g wave shaping, Karplus-Strong) often used for adding harmonics or distortion to simple sound signals~\cite{smith1991viewpoints}.

Synthesizers are engines of synthesis that make use of one or more of these techniques for sound generation. Selection of the appropriate synthesis methods depends not only on the expectations from a single end product, but also the features of the synthesizer itself. Whether in goal oriented tasks such as text-to-speech or in creative endeavors such as ambient-noise generation, it is often desirable to work with systems that are quick, adaptable and tractable. For example, one might desire a text-to-speech system where slight changes to input parameters can introduces slight changes to the speech patterns, utterances, voices etc. This ability to quickly modify and audition sounds becomes a necessity when the synthesizer is being used as a creative instrument of itself, rather than an emulator for existing instruments and sounds.

In our work, we require our synthesizer to be simple and efficient, as we would like to deploy as many instances as possible (e.g when searching for particular sounds). We also value replicability as we would like to save and re-use synthesizer programs which we found interesting. Another requirement is tractability as we would like modify parameters of a program with various levels of intensity and contrast the audition results.

Having considered our requirements, we opted to mainly utilize \enquote{additive} and \enquote{subtractive} synthesis, umbrella terms for some of the most simple and common methods of synthesis~\cite{mitchell2009basicsynthChap1}. 
In additive synthesis, sounds are built as a sum of signals; Where signals are outputs of periodic oscillators.  In subtractive synthesis, segments of a complex signal are removed until a desired sound is reached, often by a chain of one or more \textit{digital filters}. It is not uncommon for percussive sounds to have noisy, chaotic high frequency content during their short attack envelope followed by harmonic low/medium frequency content~\cite{lakatos2000common}. We hypothesize that additive synthesis will be an appropriate when precisely tuned oscillators are needed, and subtractive synthesis can be utilized for sculpting white noise to recreate chaotic, noisy segments of the spectrum.

\subsection{Filters, Oscillators ...}
[quick overview of filters and oscillators]
[will use spectrums to describe these]

\subsection{Analog Synthesizer}
Talk about analog synthesis briefly, so the word synthesizer feels less abstract...
\subsection{Virtual Synthesizers}
5 decades ago, Mathews claimed that any sound can be recreated via a computer by high frequency sampling of pressure waves. Additionally, he added that since \enquote{a very high sampling rate is required, and, if this process is to be useful musically, programs for generating samples from the parameters of notes must be written}~\cite{mathews1963digital}. The methods of synthesis we discussed in Section~\ref{sec_digital_synthesis} are a major component of such programs. With the exception of physical synthesis, modern chips are more than capable of simultaneously running many instances of these algorithms. To further assist with their musical utility, the majority of digital synthesis systems work in tandem with programs such as Musical Instrument Digital Interface (MIDI), which can  modulating the parameters of these synthesis methods by modulating information pretaining to ADSR and other note characteristics, often in real time~\cite{moog1986midi}.

 The rise of Digital Audio Workstations (DAW) ~\cite{leider2004digital} and Virtual Studio Technology (VST) based plug-ins~\cite{tanev2013virtual} have rapidly transformed the sonic and material landscape of music production in the recent years. Coupled with this rise in popularity is a vast array of commercial products and services dedicated to satiating the need of amateur and professional music producers for unique sounds; most commonly via audio samples: one-shot drum samples, long sustained notes (commonly referred to as pads or textures), and loops (percussive or melodic) are common deliverables. Two notable examples of these commercial services are \textit{loopmasters}\footnote{loopmasters.com} and \textit{splice.com}\footnote{splice.com}. Furthermore, VST plug-ins can emulate complex audio synthesizers and effects which some producers may find daunting or time consuming to work with from scratch. In many cases VST plug-in vendors or unaffiliated enthusiasts sell additional presets for these plugins, targeted towards producers who do not have the time or interest in creating their own. The flexibility of the VST technology allows producers to further modify these presets until their desired sound is reached.

\section{Neural Networks And Sound}
\label{bg:NN}
\subsection{Neural Networks For Sound Analysis}
\subsection{Neural Networks as Synthesizers}
We defined digital synthesis as the \enquote{process of generating discrete values which approximate sound waves}. We also established that manual generation of these values at high sampling rates is near impossible; A problem which has motivated a wide variety of synthesis techniques which aim to create signals within a linear (or mostly linear) system, in the simplest form possible (**look into simplest form proofs**). Recently, the exponential increase in computing power has motivated a wide range of research in probabilistic sound generation, mainly via generative neural networks. 
% deterministically, if the input parameters to the system do not change, the output of the system will remain the same \footnote{talk about noise generation methods and seeding}.

Artificial Neural networks(ANNs) map inputs to outputs via a large network of parameters. Given the right network shape and parameter weights, they can approximate a large set of functions~\cite{cybenko1989approximation,cardaliaguet1992approximation}. ANNs are often deployed when we do not have access to the system of functions which guide a process, but a mapped set of inputs and outputs of the process can be attained. Given this set, the parameters of a neural network can be tuned for approximating the effect of the system on new inputs. The shape of the neural network (e.g number of layers, connections, activation functions) is often selected via trial and error ~\cite{bergstra2012random,bergstra2011algorithms,ba2013adaptive}. By definition, these approximations will never be more accurate than the system that is being approximated.

Since their emergence in the 1950's, research on ANNs has gone through several eras of stunted growth~\cite{basheer2000artificial,anderson1988neurocomputing}. With the lack of computing power identified as a major contributor (**citations needed**). In the last decade, the increase in the affordability of high performance graphic cards has been coupled with a major resurgence of interest into ANNs. As a result, a number of domain specific variations of the traditional ANN architectures have also emerged.

Generative neural networks (GNNs) are utilized for the completion of a sequences of values; Often by taking an incomplete sequence as input and outputting the most likely value for the next step. Published in 2016 by Oord et al. WaveNet is considered a seminal breakthrough in the usage ANNs for sound synthesis~\cite{oord2016wavenet}. Surpassing state of the art concatenate and parametric speech synthesis techniques, which work by combining one or more previously recorded audio snippets~\cite{schwarz2007corpus}. When trained on a large corpus of audio samples, GNNs such WaveNet can learn the \enquote{predictive distribution for each
audio sample conditioned on all previous ones}~\cite{oord2016wavenet}. Once this distribution is learned, it can be used to create sounds 1 sample at a time, a slow process, as Mathews predicted. \\
% Parallel WaveGAN has only 1.44 M parameters and can generate 24 kHz speech waveform 28.68 times faster than real-time on a single GPU environment. Perceptual listening test results verify that our proposed method achieves 4.16 mean opinion score within a Transformer-based text-to-speech framework, which is comparative to the best distillation-based Parallel WaveNet system.
Oord et al. later proposed a parallel implementation of the WaveNet architecutre, accelerating generation by (**extract benchmarks**)~\cite{oord2017parallel}. Highly optimized GNNs such as WaveGan by Yamamoto et al.~\cite{yamamoto2020parallel} have matched WaveNet's perception performance while significantly reducing the number of parameters and speeding up synthesis (**some benchmarks in comments have to organize them**). Despite its state of the art performance, domain specific optimizations and and relatively small foot-print, WaveGAN requires the tuning of 1.44 million parameters and \enquote{generates 24 kHz speech waveform 28.68 times faster than real-time on a single GPU environment}~\cite{yamamoto2020parallel}. It should be noted that 24 kHz is half the standard quality for audio, and far lower than what is perceived as \enquote{high quality}, as previously discussed in Section~\ref{sec_sampling_rates}. We found 24 kHz and lower sampling rates to be common place in audio generation work utilizing neural networks~\cite{oord2017parallel,yamamoto2019parallel,aouameur2019neural,ramires2020neural}, likely for reducing training and generation time.

% Yet another caveat of such models is the large amount of data that is required for training as well as the hardware, energy and time to carry out the training step.

% We can create a tone via a deterministic system of equations.  Where every sample is the result of a sine function (**can i show that a sine function is the simplest possible function to create a tone?**) 
% Mathews said that we can create sound by coding in the samples one at a time ~\cite{mathews1963digital}, but this process is too slow to do by hand, which is why he advocated for digital synthesis.
% These include (i) ANNs’ success de-
% pends on both the quality and quantity of the data,
% (ii) a lack of clear rules or fixed guidelines for
% optimal ANN architecture design, (iii) a lack of
% physical concepts and relations, and (iv) the inability
% to explain in a comprehensible form the process
% through which a given decision (answer) was made
% by the ANN (i.e., ANNs criticized for being black
% boxes) ~\cite{basheer}.

% 
\section{Goal Oriented Novel Sound Generation}

>>>>

@@

\section{Goal Oriented Novel Sound Generation}
